---
title: "HW2"
author: "Shirley Liang"
date: "2022-09-26"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(readxl)

```


```{r Problem1}



```
***

### Problem 2

##### For both "Mr. Trash Wheel" and "Professor Trash Wheel" sheet: Read file and omit non-data entries, omit rows that do not include dumpster-specific data, round the number of sports balls to the nearest integer and converts the result to an integer variable

```{r Problem2 import}

importMtrashwheel <- read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>% 
  janitor::clean_names()

mtrashwheel <- mutate(importMtrashwheel, dumpster = as.numeric(dumpster))

mtrashwheel <- drop_na(mtrashwheel, dumpster) %>%
  mutate(sports_balls = round(sports_balls)) %>%
  mutate(sports_balls = as.integer(sports_balls)) 

importPtrashwheel <- read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Professor Trash Wheel", range = "A2:N117") %>% 
  janitor::clean_names()

ptrashwheel <- importPtrashwheel %>% 
  filter(!is.na(dumpster)) %>%
  mutate(sports_balls = round(sports_balls), as.integer(sports_balls)) 

```
***

##### Merge "Mr. Trash Wheel" and "Professor Trash Wheel" into "Trash Wheel" dataset.
```{r Problem2 merge}

mtrashwheel <- mutate(mtrashwheel, ID = row_number(dumpster))

ptrashwheel <- mutate(ptrashwheel, ID = row_number(dumpster)+453)

trashwheel = bind_rows(mtrashwheel, ptrashwheel)

```

##### There are `r nrow(mtrashwheel)` dumpster observations in the "Mr. Trash Wheel" sheet and `r nrow(ptrashwheel)` dumpster observations in the "Professor Trash Wheel" sheet. After merging the two sheets, there are `r nrow(trashwheel)` dumpster observations in the recording sheet "Trash Wheel". Each obseration include information about `r names(trashwheel)`. The total weight of trash collected is `r trashwheel %>% pull(weight_tons) %>% na.omit() %>% sum()`. The total number of sports balls in 2020 is `r trashwheel %>% filter(year == 2020) %>% pull(sports_balls) %>% sum()`.

### Problem 3

#### Clean pols-month.csv dataset, seprate dates, change months into text, create "president" vairable, and remove unwanted variables.
``` {r Problem3 clean dataset1}

polsmonth <- read.csv("pols-month.csv") 

polsmonth <- separate(polsmonth, col = mon, into = c('Year','Month','Day'), sep = '-') %>% mutate(polsmonth, Month = as.numeric(Month)) %>% mutate(polsmonth, Year = as.numeric(Year))

polsmonth <- polsmonth %>% mutate(Month = case_when(
  Month == 01 ~ "January", 
  Month == 02 ~ "February",
  Month == 03 ~ "March",
  Month == 04 ~ "April",
  Month == 05 ~ "May",
  Month == 06 ~ "June",
  Month == 07 ~ "July",
  Month == 08 ~ "August",
  Month == 09 ~ "Sepember",
  Month == 10 ~ "October",
  Month == 11 ~ "November",
  Month == 12 ~ "December"
  )) %>%
mutate(president = prez_dem + prez_gop) %>%
subset(select = -c(prez_dem, prez_gop, Day, mon))

```

#### Clean snp.csv dataset, seprate dates, change months into text,, and remove unwanted variables.

```{r Problem3 clean dataset2}

snp <- read.csv("snp.csv") 

snp <- separate(snp, col = date, into = c('Month','Day','Year'), sep = '/') %>% mutate(snp, Month = as.numeric(Month)) %>% mutate(snp, Year = as.numeric(Year))

snp <- snp %>% mutate(Month = case_when(
  Month == 01 ~ "January", 
  Month == 02 ~ "February",
  Month == 03 ~ "March",
  Month == 04 ~ "April",
  Month == 05 ~ "May",
  Month == 06 ~ "June",
  Month == 07 ~ "July",
  Month == 08 ~ "August",
  Month == 09 ~ "Sepember",
  Month == 10 ~ "October",
  Month == 11 ~ "November",
  Month == 12 ~ "December"
  )) %>%
subset(select = -c(Day, date))

snp <- snp %>% mutate(addyear = case_when(
  Year <= 22 ~ 2000,
  Year > 22 ~ 1900 )) %>% 
mutate(Year = Year + addyear) %>%
subset(select = -c(addyear))

```

#### Clean unemployment.csv dataset, seprate dates, change months into text,, and remove unwanted variables.

```{r Problem3 clean dataset3}

unemp <- read.csv("unemployment.csv") 

unemp <- pivot_longer(unemp, Jan:Dec, names_to = "Month", values_to = "Percentage")

unemp <- unemp %>% mutate(Month = case_when(
  Month == "Jan" ~ "January", 
  Month == "Feb" ~ "February",
  Month == "Mar" ~ "March",
  Month == "Apr" ~ "April",
  Month == "May" ~ "May",
  Month == "Jun" ~ "June",
  Month == "Jul" ~ "July",
  Month == "Aug" ~ "August",
  Month == "Sep" ~ "Sepember",
  Month == "Oct" ~ "October",
  Month == "Nov" ~ "November",
  Month == "Dec" ~ "December"
  )) 

```


#### Merge polsmonth, snp , and unemp datasets.

``` {r Problem3 Merge dataset}

mergeset <- inner_join(polsmonth, snp, by = c("Year", "Month")) %>%
            inner_join(unemp, by = c("Year", "Month"))

```





